---
title: "R Notebook"
output: html_notebook
---

hello - if you're reading this, this first section sets up what we need for the base map.


```{r}


library(osmdata)
library(smoothr)
library(tidycensus)
library(tigris)
library(tmap)
library(tmaptools)
library(units)
library(grid)
library(gridExtra)
library(ggplot2)
library(tibble)
library(stringr)
library(scales)
library(reticulate)
library(ggpubr)
library(tidyverse)
library(sf)
library(extrafont)
library(lwgeom)
library(ggspatial)
library(viridis)
library(ggrepel)

options(tigris_use_cache = TRUE)

```



NYC boundaries - honestly I have no idea what the chunk above does but run it anyway

```{r}

nyc_water <- suppressMessages(rbind(
  area_water("NY", "New York", class = "sf"),
  area_water("NY", "Kings", class = "sf"),
  area_water("NY", "Queens", class = "sf"),
  area_water("NY", "Bronx", class = "sf"),
  area_water("NY", "Richmond", class = "sf"),
  area_water("NY", "Nassau", class = "sf"),
  area_water("NY", "Westchester", class = "sf"),
  area_water("NJ", "Bergen", class = "sf"),
  area_water("NJ", "Hudson", class = "sf")) %>% 
  st_transform(26918) %>% 
  st_union())


## Import CMA, city and county boundaries
library(dplyr)
library(stringr)

nyc_msa <- suppressMessages(suppressWarnings(
  counties(state = c("New York", "New Jersey"), class = "sf") %>% 
    as_tibble() %>% 
    st_as_sf() %>% 
    st_transform(26918) %>% 
    filter(NAME %in% c("New York", "Kings", "Queens", "Bronx", "Richmond",
                       "Nassau", "Westchester", "Bergen", "Hudson", "Monmouth",
                       "Middlesex", "Somerset", "Morris", "Essex", "Union",
                       "Passaic"),
           !(STATEFP == "36" & NAME == "Essex")) %>% 
    st_erase(
      rbind(
        area_water("NY", "New York", class = "sf"),
        area_water("NY", "Kings", class = "sf"),
        area_water("NY", "Queens", class = "sf"),
        area_water("NY", "Bronx", class = "sf"),
        area_water("NY", "Richmond", class = "sf"),
        area_water("NY", "Nassau", class = "sf"),
        area_water("NY", "Westchester", class = "sf"),
        area_water("NJ", "Bergen", class = "sf"),
        area_water("NJ", "Hudson", class = "sf"),
        area_water("NJ", "Union", class = "sf"),
        area_water("NJ", "Middlesex", class = "sf"),
        area_water("NJ", "Somerset", class = "sf"),
        area_water("NJ", "Morris", class = "sf"),
        area_water("NJ", "Monmouth", class = "sf"),
        area_water("NJ", "Essex", class = "sf"),
        area_water("NJ", "Passaic", class = "sf")) %>% 
        st_transform(26918) %>% 
        st_union()
    )))
nyc_city <- nyc_msa %>%
  filter(NAME %in% c("New York", "Kings", "Queens", "Bronx", "Richmond")) %>% 
  st_union()

manhattan <- nyc_msa %>% filter(NAME == "New York")
bronx <- nyc_msa %>% filter(NAME == "Bronx")

nyc_msa <- st_union(nyc_msa)
```


```{r}
## Import PUMAs

nyc_pumas <- pumas(36, class = "sf") %>% 
  st_transform(26918) %>%
  as_tibble() %>% 
  st_as_sf() %>%
  mutate(PUMA_name = NAMELSAD10) %>% 
  select(-GEOID10, -NAMELSAD10, -STATEFP10, -MTFCC10, -FUNCSTAT10, -ALAND10,
         -AWATER10, -INTPTLAT10, -INTPTLON10) %>%
  filter(str_detect(PUMA_name, "NYC-"))
```



okay now we can create the base map!
```{r}
library(sf)
base_map <- tm_shape(nyc_msa, bbox = bb(st_bbox(nyc_city), xlim=c(-0.02, 1.02),
                                        ylim=c(0.01, 1.05), relative = TRUE),
                     unit = "mi") +
  tm_fill(col = "#f0f0f0") +
  tm_shape(nyc_city) +
  tm_fill(col = "grey80", title = "Base Map") +
  tm_scale_bar(position = c("right", "bottom"), color.dark = "grey50") +
  tm_layout(frame = TRUE, main.title.size = 1.5, legend.title.size = 1.2,
            legend.title.fontfamily = "Futura-CondensedExtraBold",
            legend.position = c("left", "top"),
            fontfamily = "Futura-Medium",
            title.fontfamily = "Futura-CondensedExtraBold")
```


this is the shape file from, pls download it: 
https://data.cityofnewyork.us/widgets/i8iw-xf4u

```{r}
nyc_zip <- st_read(dsn ="C:/Users/nour/Downloads/ZIP_CODE_040114 (1)/ZIP_CODE_040114.shp")
```

so here we will create the poulation map for each zip code, v exciting
please note if you read the nyc_zip after performing st_read, you'll note that population is one of the columns, but next we're going to mutate 1 of our datasets to add on to the shapefile, and make another map :)

```{r}
base_map +
  tm_shape(nyc_zip)  +
  tm_polygons("POPULATION", border.alpha = 0, 
              title = "Inside service area: 15.9%\nOutside service area: 20.3%", 
              palette = "-Purples",
              )

```


this is a file I made w python
dplyr is mad, but you'll get the idea in a minute
1) ETL the data you need so you have a table of zipcodes and whatever # you need
2) join with the shapefile nyc_zip
3) pass the # in 1 to tm_polygons() as before

```{r}
pop_destn = read.csv('C:/Users/nour/Downloads/most_popular_citibike_destinations.csv')
pop_destn = pop_destn %>%
  rename( zip = zip_end, num_visits = zip_end.1 )
head(pop_destn)
```

```{r}

inner_join(nyc_zip, pop_destn, by=c("ZIPCODE","zip_end"))
```